Chapter3_ideas

### 2019-12-09 results/Chapter3_ideas99.Rmd
updates to the chapter 3 analysis and the function script

# Summary and of Wang et al 2019

Main take away from Wang et al 2019 is that chromosomes with the same nucleus covary in number of crossovers. Axis lengths / SC lengths show the same pattern and is known to be regulated at the cell wide level. The authors suggest that the within nucleus covariation in CO per chromosome is mediated through cell wide regulation of SC area. (the lengths of SC for chromosomes within a cell a correlated, there is more area for COs).

Data used in the paper
human data- Gruhn et al 2013 (ploted below, 755 cells)
Sun et al 2006
Oliver-Bonet et al 2007 
Lian et al 2008
Lu et al 2012
Hou et al 2013

Was there originall CO position data for this human data? (MLH1 spreads) (there is axis length and CO count per chrm -- this should mean there is position )

@@

approach -- they compared covariation (cV) of per chrm CO count and SC length for permuted/randomized groupd of chromosome observations representing 'in silico' nuclei.

## Redoing the permutaitons

Cell summaries below

@@

@@

HS.scatter.plot <- ggplot(Hs.CO_cVs.DF, aes(y = COs, x=type) )+ geom_jitter()+ ggtitle("COs per cell, Human male")
HS.scatter.plot

#make density distributions
#percent of cells

#quantilles of COs

hist <- ggplot(Hs.CO_cVs.DF, aes(x = COs)) + geom_histogram(aes(fill=type), binwidth = 2 )
hist

@@

To calculate the variability index, authors take percent of cells(integrated curve area) which differ between the empiracal and simulated distributions.

I changed the number of permutations to 200, which should be the same number of oberved cells, but it looks like the real data is much larger....the sheet has 755 cell observations, not 200. This might be due to how histograms are presented.

The rank distinction is to test if sampling within pools of specific chromsomes makes a difference. The smaller/tighter distribution between CO count per nuc with the ranked and non-ranked distributions indecate there is a small chromosome specific effect.

Not sure how to calculate the differences in distributions yet.

this current data shows random samples, not permuted random samples.

@@

```{r setup.whole.cell.data, echo=FALSE}

#read in csv of whole cell hand measures
#setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
#setwd("C:/Users/April/Desktop/MLH1repo")
#load(file="data/MLH1/MLH1_data_setup.RData")

source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
#add sex since there is a difference in sample size
whole.cell.data <- add_mouse(whole.cell.data)
whole.cell.data <- add_sex(whole.cell.data)
whole.cell.data <- add_strain(whole.cell.data)
whole.cell.data <- add_category(whole.cell.data)

@@

#make sure foci.count is number
whole.cell.data$hand.foci.count <- as.character(whole.cell.data$hand.foci.count)
whole.cell.data$hand.foci.count <- as.numeric(whole.cell.data$hand.foci.count)

@@

#add IFDs and normalized positions 
whole.cell.data$Foci1.PER <- whole.cell.data$F1 / whole.cell.data$SC.length
whole.cell.data$Foci2.PER <- whole.cell.data$F2 / whole.cell.data$SC.length
whole.cell.data$Foci3.PER <- whole.cell.data$F3 / whole.cell.data$SC.length

@@

#Curated_BivData$IFD1_PER = Curated_BivData$IFD1_ABS / Curated_BivData$chromosomeLength
#IFD.2 does mulitple IFDs (just absolute)
#Curated_BivData <- add_IFD.2(Curated_BivData)

#find cells with the wrong number
#whole.cell.data_cell.table <-

#ADD THE RANK
for(musy in as.character(unique(whole.cell.data$fileName))){
  #print(musy)

  #order the chrm lengths -- this is order
  chrm.order = order(whole.cell.data$SC.length[whole.cell.data$fileName == musy])
  whole.cell.data$rank[whole.cell.data$fileName == musy] <- chrm.order #19 to 20
}
#I think I saved the Excel with sorted SC lengths
#Yay I printed the rank in this

@@

To make the same comparisons, I need chromosome/bivalent level observations for whole cells (all chromosomes). 
I have a data set of hand measured whole cells, in excel sheet Whole_cell_measures.xlsx with has 840 bivalent measures. The table below shows the number of bivalent observations

```{r whole.cell.table, echo=FALSE}

table(whole.cell.data$category)

#table(cell.table$category)

@@

```{r cell.level.tables, echo=FALSE}

whole.cell.data <- whole.cell.data[(whole.cell.data$hand.foci.count!="XY"), ]
male.whole.cell <- whole.cell.data[whole.cell.data$sex == "male", ]#359
#ncells
#length(unique(male.whole.cell$fileName) ) #24 cells

female.whole.cell <- whole.cell.data[whole.cell.data$sex == "female", ]#456
#length(unique(female.whole.cell$fileName) ) #18 cells

@@

                 CO = sum(hand.foci.count),
                 total.SC = sum(SC.length),
                 mean_co = mean(hand.foci.count, rm.na = TRUE),
                 mean_sc = mean(SC.length),
                   var_co = var(hand.foci.count),
                   var_sc = var(SC.length),
                 cv_co.count = cv(hand.foci.count),
                 cv_SC.length = cv(SC.length)
)

female.biv.count_table <-  ddply(.data=female.whole.cell,
                 .(fileName),
                 summarize,
                 nbiv = length(SC.length),

                  nCO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
                       nCO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
                       nCO2 = sum(hand.foci.count == 2, na.rm=TRUE),
                       nCO3 = sum(hand.foci.count == 3, na.rm=TRUE),

                 CO = sum(hand.foci.count),
                  total.SC = sum(SC.length),
                 mean_co = mean(hand.foci.count),
                 mean_sc = mean(SC.length),
                   var_co = var(hand.foci.count),
                   var_sc = var(SC.length),
                 cv_co.count = cv(hand.foci.count),
                 cv_SC.length = cv(SC.length)
)#one of the female cells is missing values

@@

I calulate within cell ranked SC length to use rank as a proxy for chromsome ID for bivalents.

@@

```{r play.plots, echo=FALSE}
#female.whole.cell
#male.whole.cell
female.whole.cell <- female.whole.cell[!(is.na(female.whole.cell$SC.length) | female.whole.cell$SC.length==""), ]

@@

rank.length.boxplot.xy <- ggplot(data=male.whole.cell, aes(x=as.factor(rank), y=SC.length)) + geom_boxplot()+ geom_jitter(aes( color=fileName)) +ggtitle("Male SC Length Dist by ranks")+theme(legend.position="none")

@@

# PERMUTATION SETUP

#
for( o in 1:nrow(DF.permutations.CO.no_rank) ) {
  DF.permutations.CO.no_rank$var_rand.co[o] <- var(as.numeric(DF.permutations.CO.no_rank[o,2:20]))
  }

#merge with real
#co.var.for.plot <- matrix(ncol=3, nrow=72)
#co.var.for.plot <- as.data.frame(co.var.for.plot)
#colnames(co.var.for.plot) <- c("cell.id","var","type")

#co.var.for.plot$var[1:24] <- DF.permutations_CO.count$var_CO.count
#co.var.for.plot$type[1:24] <- "sim"

#co.var.for.plot$var[25:48] <- male.biv.count_table$var_co
#co.var.for.plot$type[25:48] <- "real"

#co.var.for.plot$var[49:72] <- DF.permutations.CO.no_rank$var_rand.co
#co.var.for.plot$type[49:72] <- "sim_no_rank"

#plot
#whole.cell.var_CO.plot  <- ggplot(co.var.for.plot, aes(y=var, x=type))+geom_jitter()+ggtitle("within cell CO variance")
#whole.cell.var_CO.plot

@@

Make list of DFs to fill with permutations

```{r list.of.DF, echo=FALSE}

#dataframes to recreate -- and fill with 
#24 rows (cells/in silico nuclei) by 19 cols, (bivalent traits)

#rank (sample all of the ranks then stich them together)
#replication should match the num
BIG.LIST_w.rank <- replicate(18,
  list(bind_rows(list( 
  male.whole.cell[ sample( which(male.whole.cell$rank==1), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==2), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==3), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==4), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==5), 1, replace = FALSE ),],
   male.whole.cell[ sample( which(male.whole.cell$rank==6), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==7), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==8), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==9), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==10), 1, replace = FALSE ),],

@@

#LIST of DF, no rank (include all), samples all 19 bivs at a time
#makes a list of 10 -- [[]]
BIG.DF.LIST_no.rank <- replicate(18,
                  list(bind_rows(list( 
          male.whole.cell[ sample( which( male.whole.cell$rank < 20), 19, replace = FALSE ),]
) ) ) )
```

```{r sum.stat, echo=FALSE}
#dealing with lists of DFs, 
#2. #Loop through the list of DFs and do the ... make the cell level cVs calculations
#i is this index of the dataframes in the list?

@@

#then add the real values
real.vals.cv.co <- data.frame(male.biv.count_table$cv_co.count,male.biv.count_table$CO )
real.vals.cv.co$type <- "real"
colnames(real.vals.cv.co) <- c("cV","CO", "type")

@@

```{r perm.plots, echo=FALSE}
#plots for permutations

@@

Above plot shows overdispersion of CO counts per cell for my Mus male dataset.
The above plot the real data might have more variance because it's a combination of strains

@@

```{r rand.DF_CO.count.female, echo=FALSE}

#dataframes to recreate -- and fill with 
#24 rows (cells/in silico nuclei) by 19 cols, (bivalent traits)

@@

These permuted dataframes can be reused for all the meiotic traits, (CO.count, SC length, and position).

```{r sum.stat_cv.CO, echo=FALSE}
#CALUCULATING SUMMARY STAT from the permuted DFs
#dealing with lists of DFs, 
#2. #Loop through the list of DFs and do the ... make the cell level cVs calculations
#i is this index of the dataframes in the list?

@@

#then add the real values
real.vals.cv.co_female <- data.frame(cv = female.biv.count_table$cv_co.count, CO = female.biv.count_table$CO)
real.vals.cv.co_female$type <- "real"
colnames(real.vals.cv.co_female) <- c("cV","CO", "type")

@@

```{r female.plot, echo=FALSE}

Density plot above also shows that the female CO count data is overdispersed compered to a randomized distribution.

@@

Start doing the same thing but for bivalent SC lengths.
Use the same randomized data sets!
summary stats - mean SC length, cV SC length, total SC area

don't need to remake permuted DFs, reuse the DFs:  BIG.LIST_w.rank, BIG.DF.LIST_no.rank, BIG.LIST_w.rank_female and BIG.DF.LIST_no.rank_female.

```{r SC_Sum.Stats.male, echo=FALSE}
# calq the sum stats
SC_Mus.male_w.rank <- data.frame(  SC.cv = rep(0, length(BIG.LIST_w.rank) ) ) #first initialize the DF results will go in outside of loop
for(j in 1:length(BIG.LIST_w.rank)){  #inside loop, move thru the LISTs of DFs
 # print(j)
  #fill in the cells of given index
  SC_Mus.male_w.rank$SC.cv[j] <-cv(  BIG.LIST_w.rank[[j]]$SC.length ) #for each
  SC_Mus.male_w.rank$total_SC[j] <-sum(  BIG.LIST_w.rank[[j]]$SC.length )
  SC_Mus.male_w.rank$mean_SC[j] <-mean(  BIG.LIST_w.rank[[j]]$SC.length )
  }

@@

```{r SC.plots_Mus.male, echo=FALSE}
Mus_SC_hist <- ggplot(SC_Mus_male, aes(x = tot.SC)) + geom_histogram(aes(fill=type), binwidth = 50 )+ggtitle("total SC per cell, Mus male")

@@

```{r SC_sum.stats.female, echo=FALSE, eval=FALSE}

@@

#Calq cV for no rank
SC_Mus.female_no.rank <- data.frame(  cV = rep(0, length(BIG.DF.LIST_no.rank_female.) ) ) #first initialize the DF results will go in outside of loop
for(i in 1:length(BIG.DF.LIST_no.rank)){  #inside loop, move thru the LISTs of DFs
  #print(i)
  #fill in the cells of given index
  SC_Mus.female_no.rank$cV[i] <-cv(  BIG.DF.LIST_no.rank_female.[[i]]$SC.length )
  SC_Mus.female_no.rank$total_SC[j] <-sum(  BIG.DF.LIST_no.rank_female.[[j]]$SC.length )
  SC_Mus.female_no.rank$mean_SC[j] <-mean(  BIG.DF.LIST_no.rank_female.[[j]]$SC.length )
}

@@

```{r permuted.v.real, echo=FALSE, eval=FALSE}
for( o in 1:nrow(DF.permutations_CO.count) ) {
  DF.permutations_CO.count$var_CO.count[o] <- var(as.numeric(DF.permutations_CO.count[o,2:20]))
  }

#two data frames of within cell var
#cell.id, var, type
var.for.plot <- matrix(ncol=3, nrow=48)
var.for.plot <- as.data.frame(var.for.plot)
colnames(var.for.plot) <- c("cell.id","var","type")

var.for.plot$var[1:24] <- DF.permutations_CO.count$var_CO.count
var.for.plot$type[1:24] <- "sim"

var.for.plot$var[24:48] <- male.biv.count_table$var_co
var.for.plot$type[24:48] <- "real"

#my permuted DFs are in a strange str

#plot within cell variance
whole.cell.var.plot  <- ggplot(var.for.plot, aes(y=var, x=type))+geom_jitter()+ggtitle("within cell variance between types")

whole.cell.var.plot
```

@@

don't need to remake permuted DFs, reuse the DFs:  BIG.LIST_w.rank, BIG.DF.LIST_no.rank, BIG.LIST_w.rank_female and BIG.DF.LIST_no.rank_female.

look at CO position metrics.
Do I need to melt the plots 
the position col names are also, messed up I need to fix them --

1. use the same randomized dataframe
2. use the summary stat
3. draw rand nrm foci position ... then map that onto the raw SC length

How do I randomize CO pos for a chrm?
I want to draw positions from chrm_classes

For each cell observation, note the classes,  I want to 
make bins for positions of all

@@

#make a copy of main DF
male.whole.cell.mini <- male.whole.cell#[190:455, ]

for(oar in 1:length(male.whole.cell.mini$fileName)){
 # print(oar)
   if(male.whole.cell.mini$hand.foci.count[oar] == 1){
    print("oner")
     #random sample -- there could be replicates in the random sample
     rand.obj.id.1CO <- sample(pos.pool_1CO$Obj.ID, 1)

@@

### merge daata

@@

@@

### plot relative foci positions

```{r Pos.plot, echo=FALSE}

#density or histogram plot facet hand foci count, plot value, fill type

pos.plot <- ggplot(Position.DF[Position.DF$hand.foci.count == 1,], aes(x=as.numeric(value), fill=type))+geom_density(alpha=.6)

@@

pos.plot.all <- ggplot(Position.DF, aes(x=as.numeric(value), fill=type))+geom_density(alpha=.6)+facet_wrap(~hand.foci.count,ncol = 1)+ggtitle("Simulated and normalized positions, male mus")

pos.plot.all

>>>>

# Within Nuc covariation

Main take away from Wang et al 2019 is that chromosomes with the same nucleus co-vary in number of crossovers. Axis lengths / SC lengths show the same pattern and is known to be regulated at the cell wide level. The authors suggest that the within nucleus co-variation in CO per chromosome is mediated through cell wide regulation of SC area. (the lengths of SC for chromosomes within a cell a correlated, there is more area for COs).

Data used in the paper
human data- Gruhn et al 2013 (plotted below, 755 cells)
Sun et al 2006
Oliver-Bonet et al 2007 
Lian et al 2008
Lu et al 2012
Hou et al 2013

Was there original CO position data for this human data? (MLH1 spreads) (there is axis length and CO count per chrm -- this should mean there is position )

@@

## Permutations of Human Data

Heads of the cell-wide summary statistics

@@

I ran two types of permutations for the CO count per cell data. 'sim no rank' is a dataset where chromosome CO counts were drawn from the entire pool of chromosome observations. 'Sim w.rank' is a randomized pool, but each cell is made of random draws from one of each of the chromosome bins.

I didn't test the axis length trait yet.

@@

HS.scatter.plot <- ggplot(Hs.CO_cVs.DF, aes(y = COs, x=type) )+ geom_jitter()+ ggtitle("COs per cell, Human male")

hist <- ggplot(Hs.CO_cVs.DF, aes(x = COs)) + geom_histogram(aes(fill=type), binwidth = 2 )

@@

Qualitatively these results look the same as in the Wang et al 2019 paper. I haven't confirmed the calculations with a variability index. To calculate the 'variability index', the authors take the percentage of cells (or integrated curve area) which differ between the empirical and simulated distributions. I still need to do some work to make sure I can calculate the differences in distributions.

The rank distinction is to test if sampling within pools of specific chromosomes makes a difference. The smaller/tighter distribution between CO count per nuc with the ranked and non-ranked distributions indicate there is a small chromosome specific effect.

@@

```{r setup.whole.cell.data, echo=FALSE, warning=FALSE}

#add sex since there is a difference in sample size
whole.cell.data <- add_mouse(whole.cell.data)
whole.cell.data <- add_sex(whole.cell.data)
whole.cell.data <- add_strain(whole.cell.data)
whole.cell.data <- add_category(whole.cell.data)

@@

whole.cell.data$hand.foci.count <- as.numeric(as.character(whole.cell.data$hand.foci.count))
whole.cell.data$SC.length <- as.numeric(as.character(whole.cell.data$SC.length))

whole.cell.data$F1 <- as.numeric(as.character(whole.cell.data$F1))
whole.cell.data$F2 <- as.numeric(as.character(whole.cell.data$F2))
whole.cell.data$F3 <- as.numeric(as.character(whole.cell.data$F3))

@@

#add IFDs and normalized positions 
whole.cell.data$Foci1.PER <- whole.cell.data$Foci1 / whole.cell.data$SC.length
whole.cell.data$Foci2.PER <- whole.cell.data$Foci2 / whole.cell.data$SC.length
whole.cell.data$Foci3.PER <- whole.cell.data$Foci3 / whole.cell.data$SC.length

@@

#ADD THE RANK -- this stoped working for a bit, 
#fixed it with rank
for(musy in as.character(unique(whole.cell.data$fileName))){
  #print(musy)
  #order the chrm lengths -- this is order
  chrm.order = rank(whole.cell.data$SC.length[whole.cell.data$fileName == musy])
  whole.cell.data$rank[whole.cell.data$fileName == musy] <- chrm.order   #this is just a string of numbers--

@@

I have a data set of hand measured whole cells, in excel sheet Whole_cell_measures.xlsx with has 840 bivalent measures. The table below shows the number of bivalent observations by categories.

The whole cell comparisons (from empirical and randomized distributions):
- CO counts per cell
- Total SC per cell (calculated by single bivalents)
- CO position (cell wide rbar, mean telomere bias)

The CO position metrics might change.

```{r whole.cell.table, echo=FALSE}
table(whole.cell.data$category)
#table(cell.table$category)

@@

```{r make.nother.table, echo=FALSE}

whole.cell.biv.count_table <-  ddply(.data=whole.cell.data,
                 .(fileName),
                 summarize,
                 nbiv = length(SC.length),
                  #THIS IS WHERE THE error is thrown
                 nCO0 =  sum(which(hand.foci.count == 0), na.rm=TRUE)
                 #SUM not meaningful for factors
)

                 #  nCO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
            #      nCO1 =  sum(hand.foci.count == 1, na.rm=TRUE),# 
             #     nCO2 = sum(hand.foci.count == 2, na.rm=TRUE),
            #      nCO3 = sum(hand.foci.count == 3, na.rm=TRUE),

          #       totCO = sum(hand.foci.count),
          #       total.SC = sum(SC.length),
          #       mean_co = mean(hand.foci.count, na.rm = TRUE),
          #       mean_sc = mean(SC.length),
              #     var_co = var(hand.foci.count),
              #     var_sc = var(SC.length),
              #   cv_co.count = cv(hand.foci.count),
             #    cv_SC.length = cv(SC.length)
#)

```{r cell.level.tables, echo=FALSE}
whole.cell.data <- whole.cell.data[(whole.cell.data$hand.foci.count!="XY"), ]

male.whole.cell <- whole.cell.data[whole.cell.data$sex == "male", ]#359

female.whole.cell <- whole.cell.data[whole.cell.data$sex == "female", ]#456

@@

                 totCO = sum(hand.foci.count),
                 total.SC = sum(SC.length),
                 mean_co = mean(hand.foci.count, na.rm = TRUE),
                 mean_sc = mean(SC.length),
                   var_co = var(hand.foci.count),
                   var_sc = var(SC.length),
                 cv_co.count = cv(hand.foci.count),
                 cv_SC.length = cv(SC.length)
)

female.biv.count_table <-  ddply(.data=female.whole.cell,
                 .(fileName),
                 summarize,
                 nbiv = length(SC.length),
                 nCO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
                nCO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
                 nCO2 = sum(hand.foci.count == 2, na.rm=TRUE),
                 nCO3 = sum(hand.foci.count == 3, na.rm=TRUE),

                 totCO = sum(hand.foci.count, na.rm = TRUE),
                  total.SC = sum(SC.length, na.rm = TRUE),
                 mean_co = mean(hand.foci.count, na.rm = TRUE),
                 mean_sc = mean(SC.length, na.rm = TRUE),
                   var_co = var(hand.foci.count),
                   var_sc = var(SC.length),
                 cv_co.count = cv(hand.foci.count),
                 cv_SC.length = cv(SC.length)
)

@@

I calculate within cell ranked SC length to use rank as a proxy for chromosome ID for bivalents.

@@

```{r play.plots, echo=FALSE, warning=FALSE}
#female.whole.cell
#male.whole.cell
female.whole.cell <- female.whole.cell[!(is.na(female.whole.cell$SC.length) | female.whole.cell$SC.length==""), ]

@@

rank.length.boxplot.xy <- ggplot(data=male.whole.cell, aes(x=as.factor(rank), y=as.numeric(SC.length) )) + geom_boxplot()+ geom_jitter(aes( color=fileName)) +ggtitle("Male SC Length Dist by ranks")+theme(legend.position="none")

@@

The plots above look much different than when I first made them. The SC values might not be right.

# Mus Permutation Setup

I will make the permuted data sets once and display the summary stats (and compare the empirical to random distribution).

I will us the same rank and non-ranked version of permutations as I did for the human data. Since I don't have chromosome IDs, I will use the within cell SC length ranks as a proxy for chromosome identity (rank).

@@

Theses code chunks make a data set of 24 nuclei with 19 random chromosome draws each.

```{r list.of.DF, echo=FALSE, warning=FALSE}
#rank (sample all of the ranks then stich them together)
#replication should match the num
BIG.LIST_w.rank <- replicate(24,
  list(bind_rows(list( 
  male.whole.cell[ sample( which(male.whole.cell$rank==1), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==2), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==3), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==4), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==5), 1, replace = FALSE ),],
   male.whole.cell[ sample( which(male.whole.cell$rank==6), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==7), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==8), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==9), 1, replace = FALSE ),],
  male.whole.cell[ sample( which(male.whole.cell$rank==10), 1, replace = FALSE ),],

@@

#LIST of DF, no rank (include all), samples all 19 bivs at a time
#makes a list of 10 -- [[]]
BIG.DF.LIST_no.rank <- replicate(24,
                  list(bind_rows(list( 
          male.whole.cell[ sample( which( male.whole.cell$rank < 20), 19, replace = FALSE ),]
) ) ) )
```

```{r sum.stat, echo=FALSE, warning=FALSE}
#dealing with lists of DFs, 
#2. #Loop through the list of DFs and do the ... make the cell level cVs calculations
#i is this index of the dataframes in the list?

@@

#then add the real values
real.vals.cv.co <- data.frame(male.biv.count_table$cv_co.count, male.biv.count_table$totCO )
real.vals.cv.co$type <- "real"
colnames(real.vals.cv.co) <- c("cV","CO", "type")

@@

#  geom_histogram(bins = 50)+facet_wrap(.~Test.type, scales = "free") 
#biggie <- ggplot(Melted.pval.DF, aes(pval) )+geom_histogram(bins = 50)+facet_wrap(.~Test.type, scales = "free") + ggtitle("pvalues for 1000 replicated t.tests")
```

```{r perm.plots, echo=FALSE, warning=FALSE}
#plots for permutations

@@

Above plot shows over dispersion of CO counts per cell for my Mus male data set.
The above plot the real data might have more variance because it's a combination of strains

@@

These code chunks make a randomized data set of 18 nuclei with 20 random chromosome draws each.

```{r rand.DF_CO.count.female, echo=FALSE, warning=FALSE}

@@

```{r sum.stat_cv.CO, echo=FALSE, warning=FALSE}
#CALUCULATING SUMMARY STAT from the permuted DFs
#dealing with lists of DFs, 
#2. #Loop through the list of DFs and do the ... make the cell level cVs calculations
#i is this index of the dataframes in the list?

@@

#then add the real values
real.vals.cv.co_female <- data.frame(cv = female.biv.count_table$cv_co.count, CO = female.biv.count_table$totCO)
real.vals.cv.co_female$type <- "real"
colnames(real.vals.cv.co_female) <- c("cV","CO", "type")

@@

Density plot above also shows that the female CO count data is over dispersed compered to a randomized distribution.

@@

Comparing cell wide statistics for SC lengths across the real and randomized data sets.

```{r SC_Sum.Stats.male, echo=FALSE, warning=FALSE}
# calq the sum stats
SC_Mus.male_w.rank <- data.frame(  SC.cv = rep(0, length(BIG.LIST_w.rank) ) ) #first initialize the DF results will go in outside of loop
for(j in 1:length(BIG.LIST_w.rank)){  #inside loop, move thru the LISTs of DFs
 # print(j)
  #fill in the cells of given index
  SC_Mus.male_w.rank$SC.cv[j] <-cv(  BIG.LIST_w.rank[[j]]$SC.length ) #for each
  SC_Mus.male_w.rank$total_SC[j] <-sum(  BIG.LIST_w.rank[[j]]$SC.length )
  SC_Mus.male_w.rank$mean_SC[j] <-mean(  BIG.LIST_w.rank[[j]]$SC.length )
  }

@@

```{r SC.plots_Mus.male, echo=FALSE, warning=FALSE}
Mus_SC_hist <- ggplot(SC_Mus_male, aes(x = tot.SC)) + geom_histogram(aes(fill=type), binwidth = 50 )+ggtitle("total SC per cell, Mus male")

@@

```{r SC_sum.stats.female, echo=FALSE}

@@

#Calq cV for no rank,   BIG.DF.LIST_no.rank_female
SC_Mus.female_no.rank <- data.frame(  cV = rep(0, length(BIG.DF.LIST_no.rank_female) ) ) #first initialize the DF results will go in outside of loop
for(i in 1:length(BIG.DF.LIST_no.rank_female)){  #inside loop, move thru the LISTs of DFs
  #print(i)
  #fill in the cells of given index
  SC_Mus.female_no.rank$cV[i] <-cv(  BIG.DF.LIST_no.rank_female[[i]]$SC.length )
  SC_Mus.female_no.rank$total_SC[i] <-sum(  BIG.DF.LIST_no.rank_female[[i]]$SC.length, na.rm = TRUE )
  SC_Mus.female_no.rank$mean_SC[i] <-mean(  BIG.DF.LIST_no.rank_female[[i]]$SC.length, na.rm=TRUE )
}

#10 and 10 rows from the permuted
SC_Mus.female_w.rank$type <- "sim w.rank"
colnames(SC_Mus.female_w.rank) <- c("SC.cV", "tot.SC", "mean.SC", "type" )

#this DF is empty
SC_Mus.female_no.rank$type <- "sim_no.rank"
colnames(SC_Mus.female_no.rank) <- c("SC.cV", "tot.SC", "mean.SC", "type" )

real_SC.Mus_female <- data.frame(female.biv.count_table$var_sc,  female.biv.count_table$total.SC, female.biv.count_table$mean_sc )
real_SC.Mus_female$type <- "real"

colnames(real_SC.Mus_female) <- c("SC.cV", "tot.SC", "mean.SC", "type")
#merge
SC_Mus_female <- rbind(SC_Mus.female_w.rank, SC_Mus.female_no.rank, real_SC.Mus_female)

Mus_SC_density_female <- ggplot(SC_Mus_female, aes(x = tot.SC, fill=type) )+geom_density(alpha=.6) + ggtitle("Total SC per cell, Mus female")

Mus_SC_density_female

@@

The distribution of the female SC lengths isn't normal. Double check these values -- if their hand measured, not sure why 1 has much lower sc lengths.

I think the difference in variance of the rank and non-rank suimulations shows that chromsomes effects play a bigger role in shaping the distributions of cell wide SC area.

@@

I am still working on the metrics to compare (and the ones that are best for cell wide descriptions).
- cell wide telomere bias
- cell wide rbar (smaller values mean more equal chrm proportions between foci. )
- average IFD

(should these metrics should be weighted by CO count?)

## Male CO Position

Comparing cell wide statistics for CO position metrics across the real and randomized data sets.

The permuted/randomized data set is made by keeping CO count and SC lengths constant, but replacing the normalized foci position with a random normalized foci position(s) drawn from a pool of the same CO counts. The randomized relative positions are converted into raw/absolute positions (for the original SC length of that bivalent observation).

@@

#make a copy of main DF
male.whole.cell.mini <- male.whole.cell

#filling in with randomized foci positions
for(oar in 1:length(male.whole.cell.mini$fileName)){
 # print(oar)
   if(male.whole.cell.mini$hand.foci.count[oar] == 1){
    #print("oner")
     #random sample -- there could be replicates in the random sample
     rand.obj.id.1CO <- sample(pos.pool_1CO$Obj.ID, 1)

@@

### merge data

@@

male.whole.cell.mini$SC.length <- as.numeric(as.character(male.whole.cell.mini$SC.length))
male.whole.cell.mini$rand.Foci2 <- as.numeric(as.character(male.whole.cell.mini$rand.Foci2))
male.whole.cell.mini$rand.Foci3 <- as.numeric(as.character(male.whole.cell.mini$rand.Foci3))

male.whole.cell.mini$raw.rand_Foci1 <- male.whole.cell.mini$rand.Foci1 * male.whole.cell.mini$SC.length
male.whole.cell.mini$raw.rand_Foci2 <- male.whole.cell.mini$rand.Foci2 * male.whole.cell.mini$SC.length
male.whole.cell.mini$raw.rand_Foci3 <- male.whole.cell.mini$rand.Foci3 * male.whole.cell.mini$SC.length

@@

```{r maw.blah, echo=FALSE}
#cell.wide.rbar
#calq telo distnce
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")

#make a new DF
#
real_rbar.Mus <- data.frame(male.whole.cell.mini$rbar )
real_rbar.Mus$type <- "real"
colnames(real_rbar.Mus) <- c("rbar", "type" )

rand_rbar.Mus <- data.frame(male.whole.cell.mini$rand.rbar )
rand_rbar.Mus$type <- "rand"
colnames(rand_rbar.Mus) <- c("rbar",  "type" )

#merge
Mus_rbar_male <- rbind(rand_rbar.Mus, real_rbar.Mus)

#
density.plot_rbars <- ggplot(Mus_rbar_male, aes(x = rbar, fill=type) )+geom_density(alpha=.6) + ggtitle("Bivalent Rbar distributions, Mus male")+xlim(c(0,.3))
density.plot_rbars

mapp.cell.wide_cells <-  ddply(.data=male.whole.cell.mini,
                 .(fileName),
                 summarize,
                 nbiv = length(SC.length),
                 tot.SC = sum(SC.length),

                mean_rbar = mean(rbar),
                mean_rand.rbar = mean(rand.rbar),
                sum_rbar = sum(rbar),
                sum_rand.rbar = sum(rand.rbar)

)

real_cellwide_rbar.Mus <- data.frame(mapp.cell.wide_cells$mean_rbar )
real_cellwide_rbar.Mus$type <- "real"
colnames(real_cellwide_rbar.Mus) <- c("rbar", "type" )

rand_cellwide_rbar.Mus <- data.frame(mapp.cell.wide_cells$mean_rand.rbar )
rand_cellwide_rbar.Mus$type <- "rand"
colnames(rand_cellwide_rbar.Mus) <- c("rbar",  "type" )

#merge
cell.wide.Mus_rbar_male <- rbind(real_cellwide_rbar.Mus, rand_cellwide_rbar.Mus)

#
density.plot_cell.wide.rbars <- ggplot(cell.wide.Mus_rbar_male, aes(x = rbar, fill=type) )+geom_density(alpha=.6) + ggtitle("Cell wide mean Rbar distributions, Mus male")+xlim(c(0,.3))
density.plot_cell.wide.rbars

real_tot.cell_rbar.Mus <- data.frame(mapp.cell.wide_cells$sum_rbar )
real_tot.cell_rbar.Mus$type <- "real"
colnames(real_tot.cell_rbar.Mus) <- c("sum.rbar", "type" )

rand_tot.cell_rbar.Mus <- data.frame(mapp.cell.wide_cells$sum_rand.rbar )
rand_tot.cell_rbar.Mus$type <- "rand"
colnames(rand_tot.cell_rbar.Mus) <- c("sum.rbar",  "type" )
#merge
cell.wide.tot.rbar_Musmale <- rbind(real_tot.cell_rbar.Mus, rand_tot.cell_rbar.Mus)

#
density.plot_cell.wide.rbars <- ggplot(cell.wide.tot.rbar_Musmale, aes(x = sum.rbar, fill=type) )+geom_density(alpha=.6) + ggtitle("Cell wide total Rbar distributions, Mus male")+xlim(c(0,4))

density.plot_cell.wide.rbars

The random relative foci positions were 'mapped' onto each of the SC length observations (foci * SC.length).

The first plot shows the 2 disttributions of all the rbar measures (across all bivalents)
The second plot shows the mean rbar for the whole cell

```{r relative.foci.Pos.plot, echo=FALSE}

#mapp.cell.wide_cells
#density or histogram plot facet hand foci count, plot value, fill type
pos.plot <- ggplot(Position.DF[Position.DF$hand.foci.count == 1,], aes(x=as.numeric(value), fill=type))+geom_density(alpha=.6)

@@

pos.plot.all <- ggplot(Position.DF, aes(x=as.numeric(value), fill=type))+geom_density(alpha=.6)+facet_wrap(~hand.foci.count,ncol = 1)+ggtitle("Simulated and real normalized positions, male mus")
pos.plot.all

pos.plot.one <- ggplot(Position.DF[Position.DF$hand.foci.count == 2,], aes(x=as.numeric(value), fill=type))+geom_density(alpha=.6)+facet_wrap(~hand.foci.count)+ggtitle("Simulated and real normalized positions")
```
